
% Plain TeX, not LaTeX.

% ------------------------------------------------------------------------------
% Setup
% ------------------------------------------------------------------------------

% ------------------------
% basics
% ------------------------

\let\noarrow  = t
\let\nobibtex = t
\ifx\eplain\undefined \input eplain \fi

\magnification = \magstep0
\baselineskip  = 12pt
\parskip       = 0.5\baselineskip
\parindent     = 0.0truein

\raggedbottom
\leftdisplays
\nopagenumbers

\special{papersize=8.5in,11.0in}
\hsize   = 6.5truein
\vsize   = 9.0truein
\hoffset = 0.0truein
\voffset = 0.0truein

% suppress "underfull \hbox" messages
\hbadness=10000


% ------------------------
% fonts
% ------------------------

\font\hvbig = fvsb8r  at 16pt
\font\hv    = fvsr8r  at 10pt
\font\bf    = fvsb8r  at 10pt
\font\bfmed = fvsb8r  at 11pt
\font\it    = fvsro8r at 10pt
\font\tt    = fvmr8r  at 10pt
\font\ttbf  = fvmb8r  at 10pt
\hv


% ------------------------
% macros
% ------------------------

\def\section#1{\bigskip{\bfmed#1.}}



% ------------------------------------------------------------------------------
% Title
% ------------------------------------------------------------------------------

\centerline{\hvbig Ristra Atomics Library}
\medskip
\smallskip



% ------------------------------------------------------------------------------
% Background
% ------------------------------------------------------------------------------

\section{Background}
The Standard C++ Library's {\tt <atomic>} header defines an {\tt atomic<T>}
class template with basic functionality for any trivially-copyable type {\tt T};
informally, any type {\tt T} for which a shallow copy, as with {\tt memcpy},
has the correct semantics. Integral and pointer (but not floating-point)
specializations support {\tt \&=}, {\tt |=}, and {\tt \^{}=} (integral only);
{\tt +=} and {\tt -=}; and prefix and postfix {\tt ++} and {\tt --}. Other
viable operations, such as {\tt *=} and {\tt <<=}, are not provided.

Sandia National Laboratory's {\it Kokkos} provides a variety of atomic
operations in functional form. {\tt Kokkos::atomic\_fetch\_add(x,y)}, for
instance, is an atomic {\tt x += y}. As with C++, some viable operations
are missing. Floating-point data are supported, but pointers are not.



% ------------------------------------------------------------------------------
% Ristra Atomics
% ------------------------------------------------------------------------------

\section{Ristra Atomics}
Our new, ``header-only'' Ristra Atomics C++ library provides a class template
{\tt ristra::atomics::atomic<T,SCHEME>}, a superset of the atomic operations
of C++ and Kokkos, and several available ``atomicity schemes'' for achieving
the requisite atomicity. Schemes include our own general ones, as well as
forwards to C++ or Kokkos capabilities (Kokkos inclusion is optional) when
available. A scheme is specified by providing any of several available tag
classes as {\tt atomic}'s second template argument. This compile-time mechanism
facilitates function overloading and inlining, giving runtime efficiency.



% ------------------------------------------------------------------------------
% Atomicity Schemes
% ------------------------------------------------------------------------------

\section{Atomicity Schemes}
Ristra Atomics provides the following atomicity schemes.

{
\leftskip 0.30truein
\rightskip 0.30truein
\raggedright
\gutter = 0.01truein
\triplecolumns

\hbox{\vbox{
{\ttbf cpp}
\hfill\break
Forward to the C++ library's capability.
\hfill\break
\hfill\break
{\ttbf kokkos}
\hfill\break
Forward to the Kokkos library's capability.
\hfill\break
\hfill\break
{\ttbf lock}
\hfill\break
Embed the operation in a {\tt lock\_guard<mutex>}.
}}
\columnfill

\vskip-0.1truein % Argh
\hbox{\vbox{
{\ttbf strong}
\hfill\break
Atomicize the operation via a strong compare-and-swap (CAS) loop.
\hfill\break
\hfill\break
{\ttbf strong::pun}
\hfill\break
Like {\tt strong}, but with {\tt T} punned to a same-sized integral type
for CAS. May give better performance on some architectures.
}}
\columnfill

\hbox{\vbox{
{\ttbf weak}
\hfill\break
Like {\tt strong}, but using a weak CAS.
\hfill\break
\hfill\break
{\ttbf weak::pun}
\hfill\break
Like {\tt strong::pun}, but using a weak CAS.
\hfill\break
\hfill\break
{\ttbf serial}
\hfill\break
Perform the operation without atomicizing it.
}}
\columnfill

}

\singlecolumn

% Another awesome TeXism.
% Because NOT doing incomprehensible formatting stuff
% would just make things too easy for the user. :-)
\vskip-3.3truein



% ------------------------------------------------------------------------------
% Operations
% ------------------------------------------------------------------------------

\section{Operations}
Atomic operations in Ristra Atomics are available in several forms.
Operators:
{\tt +=},
{\tt -=},
{\tt *=},
{\tt /=},
{\tt \%=},
{\tt <<=},
{\tt >>=},
{\tt \&=},
{\tt |=},
{\tt \^{}=},
{\tt ++value},
{\tt value++},
{\tt --value}, and
{\tt value--}.
Functions and member functions:
{\tt add},
{\tt sub},
{\tt mul},
{\tt div},
{\tt mod},
{\tt lshift},
{\tt rshift},
{\tt andeq},
{\tt oreq},
{\tt xoreq},
{\tt preinc},
{\tt postinc},
{\tt predec}, and
{\tt postdec}.
{\bf Example}: for {\tt x} of atomic type, {\tt x.add(2)} and {\tt add(x,2)}
both perform {\tt x += 2} atomically. Also available:
{\tt inc == preinc},
{\tt dec == predec},
{\tt min}, and
{\tt max}.
{\bf Example}: {\tt x.min(y)} computes {\tt x = min(x,y)} atomically.



% ------------------------------------------------------------------------------
% Data Types
% ------------------------------------------------------------------------------

\section{Data Types}
Integral, floating-point, pointer, and general trivially-copyable data are all
supported. Bear in mind that the availability of a specific atomic computation
necessarily depends on the data type, atomicity scheme, and operation. The
{\tt kokkos} scheme doesn't support pointers, while {\tt cpp} doesn't support
floating-point values. No scheme supports bitwise operations on floating-point
values, which in C++ have no bitwise operations to begin with. Only our own
schemes support operations on values of user-defined types -- and, then, only
if the type has been outfitted with a non-atomic version of the operation.



% ------------------------------------------------------------------------------
% Summary
% ------------------------------------------------------------------------------

\section{Summary}
Ristra Atomics is an easy-to-use, header-only C++ library for performing atomic
operations on integral, pointer, floating-point, and general trivially-copyable
data. It supports several of its own atomicizing schemes, and provides access
to the capabilities in C++ and in Kokkos. The library thus brings together two
existing sets of capabilities, adds more, and embeds them all into a coherent
notational and logical interface.

\bye
